Index-aware learning of circuits

### 2024-04-05 doc/article/arxiv/article.tex
final revision

\raggedbottom
\newcommand{\mb}[1]{\mathbf{#1}}
\newcommand{\mbt}[1]{\tilde{\mathbf{#1}}}
\newcommand{\mbb}[1]{\bar{\mathbf{#1}}}
\newcommand{\mbh}[1]{\hat{\mathbf{#1}}}
\newcommand{\mr}[1]{\mathrm{#1}}
\newcommand{\T}{{\!\top}}
\newcommand{\ddt}{\frac{\mathrm{d}}{\mathrm{d}t}}
\newcommand{\A}[1]{\mb{A}_\mr{#1}}
\newcommand{\AT}[1]{\mb{A}_\mr{#1}^{\T}}
\newcommand{\qC}{\mb{q}_\mr{C}}
\newcommand{\gR}{\mb{g}_\mr{R}}
\newcommand{\phiL}{\boldsymbol{\phi}_\mr{L}}
\newcommand{\vphi}{\boldsymbol{\varphi}}
\renewcommand{\i}[1]{\mb{i}_\mr{#1}}
\renewcommand{\v}[1]{\mb{v}_\mr{#1}}

@@

In order to obtain a version of \eqref{eq:md_mnai} that is better suited to analysis and implementation, we consider the device function Jacobians
\begin{align}
    \mb{G} \big( \AT{R} \vphi \big) \coloneqq \mb{J}_{\AT{R} \vphi}(\gR), \quad \mb{L}(\i{L}) \coloneqq \mb{J}_{\i{L}}(\phiL), \quad \mb{C} \big( \AT{C} \vphi \big) \coloneqq \mb{J}_{\AT{C} \vphi}(\qC), \label{eq:md_dfj}
\end{align}
where we use the same notation for the Jacobians as in \autoref{sec:int}. Inserting \eqref{eq:md_dfj} into the original system, while linearizing $\gR$ and writing everything in matrix form, yields
\begin{align}
    \begin{bmatrix}
        \A{C}^{\phantom{\T}} \mb{C} \big( \AT{C} \vphi \big) \AT{C} & &\\
        & \mb{L}(\i{L}) &\\
        & & \mb{0}
    \end{bmatrix} \ddt \begin{bmatrix}
        \vphi\\
        \i{L}\\
        \i{V}
    \end{bmatrix} + \begin{bmatrix}
        \A{R}^{\phantom{\T}} \mb{G} \big( \AT{R} \vphi \big) \AT{R} & \A{L} & \A{V}\\
        -\AT{L} & &\\
        -\AT{V} & &
    \end{bmatrix} \begin{bmatrix}
        \vphi\\
        \i{L}\\
        \i{V}
    \end{bmatrix} + \begin{bmatrix}
        \A{I} \i{s}(t)\\
        \mb{0}\\
        \v{s}(t)
    \end{bmatrix} = \mb{0}, \label{eq:md_mnaii}
\end{align}
assuming that $\phiL$ and $\qC$ do not explicitly depend on time.

>>>>

\raggedbottom
\newcommand{\mb}[1]{\mathbf{#1}}
\newcommand{\mbt}[1]{\tilde{\mathbf{#1}}}
\newcommand{\mbb}[1]{\bar{\mathbf{#1}}}
\newcommand{\mbh}[1]{\hat{\mathbf{#1}}}
\newcommand{\mr}[1]{\mathrm{#1}}
\newcommand{\T}{{\!\top}}
\newcommand{\ddt}{\frac{\mathrm{d}}{\mathrm{d}t}}
\newcommand{\A}[1]{\mb{A}_\mr{#1}}
\newcommand{\AT}[1]{\mb{A}_\mr{#1}^{\T}}
\newcommand{\qC}{\mb{q}_\mr{C}}
\newcommand{\gR}{\mb{g}_\mr{R}}
\newcommand{\phiL}{\boldsymbol{\phi}_\mr{L}}
\newcommand{\vphi}{\boldsymbol{\varphi}}
% \iL, \iV, \is, \vs
\renewcommand{\i}[1]{\mb{i}_\mr{#1}}
\renewcommand{\v}[1]{\mb{v}_\mr{#1}}

@@

In order to obtain a version of \eqref{eq:md_mnai} that is better suited to analysis and implementation, we consider the device function Jacobians\footnote{Note that the definition of $\mb{G}$ via the Jacobian only serves to obtain a matrix form in \eqref{eq:md_mnaii} and has no impact on the general approach. As such, one may also work directly with the original system from \eqref{eq:md_mnai}, however software implementations of MNA often work with $\mb{G}$ as defined in \eqref{eq:md_dfj} (e.g.~when using Newton's method for solving nonlinear systems).}
\begin{align}
    \mb{G} \big( \AT{R} \vphi \big) \coloneqq \mb{J}_{\AT{R} \vphi}(\gR), \quad \mb{L}(\i{L}) \coloneqq \mb{J}_{\i{L}}(\phiL), \quad \mb{C} \big( \AT{C} \vphi \big) \coloneqq \mb{J}_{\AT{C} \vphi}(\qC), \label{eq:md_dfj}
\end{align}
where we use the same notation for the Jacobians as in \autoref{sec:int}. Inserting \eqref{eq:md_dfj} into the original system and writing everything in matrix form yields
\begin{align}
    \begin{bmatrix}
        \A{C}^{\phantom{\T}} \mb{C} \big( \AT{C} \vphi \big) \AT{C} & &\\
        & \mb{L}(\i{L}) &\\
        & & \mb{0}
    \end{bmatrix} \ddt \begin{bmatrix}
        \vphi\\
        \i{L}\\
        \i{V}
    \end{bmatrix} + \begin{bmatrix}
        \A{R}^{\phantom{\T}} \mb{G} \big( \AT{R} \vphi \big) \AT{R} & \A{L} & \A{V}\\
        -\AT{L} & &\\
        -\AT{V} & &
    \end{bmatrix} \begin{bmatrix}
        \vphi\\
        \i{L}\\
        \i{V}
    \end{bmatrix} + \begin{bmatrix}
        \A{I} \i{s}(t)\\
        \mb{0}\\
        \v{s}(t)
    \end{bmatrix} = \mb{0}, \label{eq:md_mnaii}
\end{align}
assuming that $\phiL$ and $\qC$ do not explicitly depend on time.
